### 2023-11-08 thesis/chapters/02-analysis-B-features.tex
Journal actual work done in the plan of work
Restructure folders
Refactor feature selection methods and finish correletion ranking
Revise analysis chapter and add placeholder sections to be filled

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                    & \textbf{Feature}        & \textbf{Equation}                                                                                                  \\ \hline
\textit{Standard deviation} & $ \sigma = \sqrt{\frac{1}{N}\sum_{i = 1}^{N}{\left(x_i - \bar{x}\right)^2}} $        & \textit{Crest factor}   & $ X_{cf} = \frac{max(|x_i|)}{\left( \frac{1}{N} \sum_{i=1}^{N}{x_i^2} \right)^\frac{1}{2}} $                       \\ \hline
\textit{Skewness}           & $ X_{sv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^3}$ & \textit{Margin factor}  & $ X_{mf} = \frac{max(|x_i|)}{\left( \frac{1}{N} \sum_{i=1}^{N}{\sqrt{|x_i|}} \right)^2} $                          \\ \hline
\textit{Kurtosis}           & $ X_{kv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^4}$ & \textit{Impulse factor} & $ X_{if} = \frac{max(|x_i|)}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}} $                                                  \\ \hline
\textit{Root mean square}   & $ X_{rms} = \left(\frac{1}{N}\sum_{i = 1}^{N}{x_i^2}\right)^\frac{1}{2}$             & \textit{Shape factor}   & $ X_{sf} = \frac{\left( \frac{1}{N} \sum_{i=1}^{N}{x_i^2} \right)^\frac{1}{2}}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}}$ \\ \hline
\textit{Peak-to-peak}       & $ X_{ppv} = \max(x_i) - \min(x_i) $                                                  & \textit{Maximum}        & $ X_{max} = \max(|x_i|) $                                                                                          \\ \hline
\end{tabular}
\end{adjustbox}
\caption{Time domain features}
\label{tab:td-features}
\end{table}

@@

\myequations{MMS algorithm: Local maxima identification equality}
\begin{ceqn}\begin{align}
\mathrm{MMS}_{\mathrm{max}} = \mathrm{MMS}_{\mathrm{max|mid}} \\
\frac{a_{max} - a_{min}}{S_3 - a_{min} \cdot 3} = \frac{a_{mid} - a_{min}}{S_3 - a_{min} \cdot 3}
\label{equ:mms-maxima}
\end{align}\end{ceqn}

\myequations{MMS algorithm: Local minima identification equality}
\begin{ceqn}\begin{align}
\mathrm{MMS}_{\mathrm{min}} = \mathrm{MMS}_{\mathrm{min|mid}} \\
\frac{a_{max} - a_{min}}{a_{max} \cdot 3 - S_3} = \frac{a_{max} - a_{mid}}{a_{max} \cdot 3 - S_3}
\label{equ:mms-minima}
 \end{align}\end{ceqn}

@@

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
%\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Spectral roll-off} & $ \sum_{0}^{f_c}a^2(f) = 0.95 \sum_{0}^{f_s / 2}a^2(f)$                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \frac{\sum_k a(t-1, k) \cdot a(t,k)}{\sqrt{\sum_k a(t-1, k)^2} \sqrt{\sum_k a(t, k)^2}}$ \\ \hline 
\textit{Harmonic spectral deviation} & $ X_{\mathrm{HDEV}} = \frac{1}{H}\sum_h(a(h) - \mathrm{SE}(h))$                                          \\ \hline
\textit{Noisiness}                   & $X_{noise} = \frac{\sum_{k \notin h}^{N}E_k}{\sum_{i = 1}^{N}E_i} $                             \\ \hline
\textit{Inharmonicity}               & $X_{inharmo} = \frac{2}{f_0} \cdot \frac{\sum_h | f(h) - h \cdot f_0| \cdot a^2(h)}{\sum_h a^2(h)}$ \\ \hline
\textit{Energy}            & $ E = \sum_{i = 0}^{N - 1}|x(t)|^2 $                                                                    \\ \hline \textit{Energy ratio}                & $E_i = \frac{E_i}{\sum_{i = 1}^{N}E_i} $                                                        \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\end{tabular}
%\end{adjustbox}
\caption{Frequency domain features}
\label{tab:fd-features}
\end{table}

@@

\textbf{Empirical Wavelet Transform} (EWT) constructs adaptive bandpass filters with Meyer wavelet. The scaling function $\hat{\phi_n}$ and wavelet function are defined according to equations (\ref{equ:ewt-scaling}) and (\ref{equ:ewt-wavelet}) respectively. The inner product of the signal with the scaling function obtains the approximation coefficients and the inner product with the wavelet function results in detail coefficients.

\myequations{Empirical scaling function}
\begin{ceqn}\begin{align}
\hat{\phi_n}(\omega) = 
\begin{cases}
1       & \text{if } |\omega| \leq \omega_n - \tau_n \\
\cos\left[\frac{\pi}{2}\cdot \beta\left(\frac{1}{2\tau_n}(|\omega| - \omega_n + \tau_n)\right)\right]  &  \text{if } \omega_n  - \tau_n \leq |\omega| \leq \omega_n + \tau_n \\
0       & \text{otherwise }
\end{cases}
\label{equ:ewt-scaling}
\end{align}\end{ceqn}

\myequations{Empirical wavelet function}
\begin{ceqn}\begin{align}
\hat{\psi_n}(\omega) = 
\begin{cases}
1       & \text{if } \omega_n + \tau_n \leq |\omega| \leq \omega_{n+1} - \tau_{n+1} \\
\cos\left[\frac{\pi}{2}\cdot \beta\left(\frac{1}{2\tau_{n+1}}(|\omega| - \omega_{n+1} + \tau_{n+1})\right)\right] & \text{if } \omega_{n+1} - \tau_{n+1} \leq |\omega| \leq \omega_{n+1} + \tau_{n+1} \\
\sin\left[\frac{\pi}{2}\cdot \beta\left(\frac{1}{2\tau_n}(|\omega| - \omega_n + \tau_n)\right)\right] & \text{if }\omega_n - \tau_n \leq |\omega| \leq \omega_{n} + \tau_{n} \\
0       & \text{otherwise }
\end{cases}
\label{equ:ewt-wavelet}
\end{align}\end{ceqn}

@@

Refinements acting in the onset as preparation before wavelet processing create an envelope curve in the frequency domain and attempts to eliminate supposed interference.

@@

>>>>

% TODO: add td and fd features from tsfel
\begin{table}[h]
\centering
\renewcommand{\arraystretch}{2}
\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|l|l|l|l|}
\hline
\textbf{Feature}            & \textbf{Equation}                                                                    & \textbf{Feature}        & \textbf{Equation}                                                                                                  \\ \hline
\textit{Standard deviation} & $ \sigma = \sqrt{\frac{1}{N}\sum_{i = 1}^{N}{\left(x_i - \bar{x}\right)^2}} $        & \textit{Crest factor}   & $ X_{cf} = \frac{max(|x_i|)}{X_{rms}} $                       \\ \hline
\textit{Skewness}           & $ X_{sv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^3}$ & \textit{Margin factor}  & $ X_{mf} = \frac{max(|x_i|)}{\left( \frac{1}{N} \sum_{i=1}^{N}{\sqrt{|x_i|}} \right)^2} $                          \\ \hline
\textit{Kurtosis}           & $ X_{kv} = \frac{1}{N}\sum_{i = 1}^{N}{\left(\frac{x_i - \bar{x}}{\sigma}\right)^4}$ & \textit{Impulse factor} & $ X_{if} = \frac{max(|x_i|)}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}} $                                                  \\ \hline
\textit{Root mean square}   & $ X_{rms} = \sqrt{\frac{1}{N}\sum_{i = 1}^{N}{x_i^2}} $             & \textit{Shape factor}   & $ X_{sf} = \frac{X_{rms}}{\frac{1}{N} \sum_{i=1}^{N}{|x_i|}}$ \\ \hline
\textit{Peak-to-peak}       & $ X_{ppv} = \max(x_i) - \min(x_i) $                                                  & \textit{Maximum}        & $ X_{max} = \max(|x_i|) $                                                                                          \\ \hline
\end{tabular}
\end{adjustbox}
\caption{Time domain features}
\label{tab:td-features}
\end{table}

@@

\myequations{MMS algorithm: Local maxima identification equality}
\begin{ceqn}\begin{align}
\frac{a_{max} - a_{min}}{S_3 - a_{min} \cdot 3} = \frac{a_{mid} - a_{min}}{S_3 - a_{min} \cdot 3}
\label{equ:mms-maxima}
\end{align}\end{ceqn}

\myequations{MMS algorithm: Local minima identification equality}
\begin{ceqn}\begin{align}
\frac{a_{max} - a_{min}}{a_{max} \cdot 3 - S_3} = \frac{a_{max} - a_{mid}}{a_{max} \cdot 3 - S_3}
\label{equ:mms-minima}
 \end{align}\end{ceqn}

@@

\begin{table}[ht]
\renewcommand{\arraystretch}{2}
\centering
%\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|l|l|}
\hline
\textbf{Feature}           & \textbf{Equation}                                                                                                  \\ \hline
\textit{Spectral centroid} & $ X_{fc} = \frac{\sum_{i = 0}^{N - 1}{f_i \cdot a(f_i)}}{\sum_{i = 0}^{N - 1}{a(f_i)}}$                   \\ \hline 
\textit{Spectral roll-off} & $ E(f_c) = 0.95 \cdot E(f_s / 2) $                                                   \\ \hline
\textit{Spectral flux}     & $X_{\mathrm{flux}} = 1 - \mathrm{corr}(A_{t-1}, A_t)$ \\ \hline 
\textit{Harmonic spectral deviation} & $ X_{\mathrm{HDEV}} = \frac{1}{H}\sum_h(a(h) - \mathrm{SE}(h))$                                          \\ \hline
\textit{Noisiness}                   & $X_{noise} = \frac{\sum_{k \notin h}^{N}E_k}{\sum_{i = 1}^{N}E_i} = \mathrm{SNR}(X) = \mu \;/\; \sigma $                             \\ \hline
\textit{Inharmonicity}               & $X_{inharmo} = \frac{2}{f_0} \cdot \frac{\sum_h | f(h) - h \cdot f_0| \cdot a^2(h)}{\sum_h a^2(h)} $ \\ \hline
\textit{Energy}            & $ E(N) = \sum_{i = 1}^{N} a^2(t) $                                                                    \\ \hline \textit{Energy ratio}                & $E_r = E_i \;/\; \sum_{i = 1}^{N}{E_i} $                                                        \\ \hline
\textit{Shannon entropy}      & $H(X) = - \sum_{i} P(X = x_i) \cdot \ln(P(X = x_i)) $                                                  \\ \hline
\end{tabular}
%\end{adjustbox}
\caption{Frequency domain features}
\label{tab:fd-features}
\end{table}

@@

\textbf{Empirical Wavelet Transform} (EWT) constructs adaptive bandpass filters with Meyer wavelet. The inner product of the signal with the scaling function $\hat{\phi_n}$ obtains the approximation coefficients and the inner product with the wavelet function results in detail coefficients.

@@

@@

% TODO: we did not utilize wavelets as of yet
% 
% 
%

### 2023-11-08 thesis/chapters/02-analysis-C-diagnostics.tex
Revise analysis chapter and add placeholder sections to be filled

@@

\subsection{Classification}
Accurate multi-class classification of machine fault causes out of the characteristics of known ones is a much more difficult task than novelty detection. Universal enough fault baseline has to be recorded and transformed into feature space. Interactions among fault manifestations have to be accounted for. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production ready. Therefore, we opt to use a more rudimentary model.
\bigbreak

@@

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, or \emph{RBF kernel} (Tab.~\ref{tab:knn-distance})~\cite{sheng_review_2020}. The demanding neighborhood queries are sped up by a Metric search tree data structure such as \emph{M-Tree} (Fig.~\ref{fig:m-tree}). The optimal $k$ parameter is set in supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|}
\hline
\textbf{Distance}     & \textbf{$d(\mathbf{x}, \mathbf{y})$}                                   \\ \hline
Euclidian distance    & $ \sqrt{\sum_{i = 1}^{n}(x_i - y_i)^2} $                               \\ \hline
Mahalanobis distance  & $ (\mathbf{x} - \mathbf{y})^T C^{-1} (\mathbf{x} - \mathbf{y}) $       \\ \hline
Radial basis function & $ \exp\left(-\frac{\lVert \mathbf{x} - \mathbf{y} \rVert^2}{t}\right) $ \\ \hline
\end{tabular}
\caption{Distance metrics for kNN}
\label{tab:knn-distance}
\end{table}

@@

>>>>

\subsection{Concept drift}
% TODO: ADWIN (ADaptive WINdowing)
% Kolmogorov-Smirnov Windowing method for concept drift detection.

@@

\subsection{Classification}
% TODO: HoeffdingTreeClassifier
% https://analyticsindiamag.com/a-beginners-guide-to-hoeffding-tree-with-python-implementation/
% https://www.cms.waikato.ac.nz/~abifet/book/chapter_5.html

Accurate multi-class classification of machine fault causes out of the characteristics of known ones is a much more difficult task than novelty detection. Universal enough fault baseline has to be recorded and transformed into feature space. Interactions among fault manifestations have to be accounted for. We are aware of rapid advances in knowledge transfer for deep neural networks~\cite{maurya_condition-based_2021}. So far, solutions seem not production ready. Therefore, we opt to use a more rudimentary model.
\bigbreak

@@

% TODO: Spatial database
% TODO: Sliding WIndow-based Nearest Neighbor (SWINN)

The sense of distance between feature vectors $\mathbf{x}$, $\mathbf{y}$ have to be defined, so several metrics are available like \emph{Euclidian distance}, \emph{Mahalanobis distance}, or \emph{Manhatann distance} (Tab.~\ref{tab:knn-distance})~\cite{sheng_review_2020}. The demanding neighborhood queries are sped up by a Metric search tree data structure such as \emph{M-Tree} (Fig.~\ref{fig:m-tree}). The optimal $k$ parameter is set in supervised learning according to the breaking point in the elbow curve that plots choices of $k$ against the error rate.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{2}
\begin{tabular}{|l|l|}
\hline
\textbf{Distance}     & \textbf{$d(\mathbf{x}, \mathbf{y})$}                                   \\ \hline
Manhatann distance	 & $ |x_i - y_i| $													   \\ \hline
Euclidian distance    & $ \sqrt{\sum_{i = 1}^{n}(x_i - y_i)^2} $                               \\ \hline
Mahalanobis distance  & $ (\mathbf{x} - \mathbf{y})^T C^{-1} (\mathbf{x} - \mathbf{y}) $       \\ \hline
\end{tabular}
\caption{Distance metrics for kNN}
\label{tab:knn-distance}
\end{table}

@@

\subsection{Online model evaluation}
% TODO

### 2023-11-08 thesis/chapters/03-design.tex
Revise analysis chapter and add placeholder sections to be filled

>>>>

\chapter{ML Pipeline Design}

Mafaulda dataset
Jupyter notebook - Python

\section{Data labeling and cleaning}
- Mafaulda labeling and creating subset (RPM range, 4 classes, Anomaly class)
- Measurement of custom datasets (Schiratech iComox, Beaglebone) - air conditioning, compressor

\section{Feature relevance}
- supervised - correlation, fisher, mi
- custom, tsfel, in sensor and in one axis
- online learning features

\section{Nearest neighbour classifier}
- kNN (fault and anomaly)
	- Offine (baseline)
	- Online (evolution)
- parameters: k-neigh, distance
- metrics: accuracy, precision, recall

\section{Clustering}
- Density based clustering
	- DBSCAN (all features, PCA, subset of best)
	- DenStream (all features, subset of best)

\section{Anomaly detection / Decision Trees}
- Offline (DecisionTree)
- Online (Hoefing Tree)

\section{Comparision to real data}
- Can be part of original chapters

### 2023-11-08 thesis/chapters/04-implementation.tex
Revise analysis chapter and add placeholder sections to be filled

\chapter{Implementation} \label{section:implementation} 
Mafaulda dataset
Jupyter notebook - Python

\section{Feature extraction}
- How looked dataframe for: time frequency, frequency domain, wavelet packet features

\section{Feature selection}
- What are the best features
- Feature normalization
- PCA and inverse transformations

\section{Fault detection models}
- Streaming algorithms: sklearn, river packages
kNN, Half-space tree (vs. Isolation forest),  DenStream (vs. DBSCAN)

>>>>

\chapter{Implementation} \label{section:implementation}
